Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-05-07 hpmoon-asoco.tex
Adds references
Revised intro
With many suggestions for @fergunet. Please check them out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\journal{Applied Soft Computing}
\title{Distributed multi-objective evolutionary optimization using island-based selective operator application}
% New title suggestion
% Should co-evolutionary go somewhere? Are they co-evolutionary? - jj
% PABLO: Renaming again to avoid using the co-evolutionary and
% cooperative
% Ok - JJ

@@

%Introduction should start in the same way as the abstract. And
%starting with something that is not really applicable here, talking
%about general parallel evolutionary algorithms, is even worse.
% Right way (IMNSHO)
% 1. Multiobjective problems are
% 2. They usually require a good amount of computation → parallel
% distributed
% 3. One way to solve them is to use MOEAs _because_ they can be
% easily parallelized
% 4. But the easy paralellization is not necessarily efficient.
% 5. We have proposed the HPMOON (or whatever) technique to improve
% it. In paper (whatever) we proved that (whatever)
% 6. In this paper we will be proving whatever - JJ FERGU: Addressing JJ order

%1. MOP ARE
Multi-objective optimization problems (MOP) are those where several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies maximizing or minimizing
a function composed of several cost functions (one per objective). In
these problems the aim is to obtain a set of solutions that are better
than the rest, considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
that is, none is better than the others for all the objectives.

%2 COMPUTATION -> PARALLEL
These kind of problems implies a high demand of computational time, leading to the apparition of easily parallelizable methods to solve them \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268}.
%3 ONE WAY IS TO USE MOEAS BECAUSE
One example of easily parallelizable algorithms to solve MOPs are the Multi-Objective Evolutionary Algorithms (MOEAs), a sub-set of Evolutionary algorithms (EAs). EAs \citep{DBLP:series/ncs/EibenS15} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regard to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met. One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies to use different distribution
and sharing mechanisms, as there exist a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}.

MOPs usually require a high number of variables, meaning that the MOEAs need to deal with large individuals and spend significant extra time for crossover, mutation and migration. Different authors have proposed methods to divide the decision space (the chromosome) to improve the performance and solution quality. In this aspect, the co-evolution model is a dimension-distributed model where a high-dimensional problem is divided into lower dimensional ones \citep{Gong15models,Tonda12cooperative}, and evolved separately. One example of application of this technique was described in \citep{Kimovski15Parallel}. The method presented in that work involves different workers that evolve sub-populations created and recombined by a master process, which performs different recombination alternatives from the parts returned by worker processes. A high dimensional problem was used to compare these
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

A similar approach used to solve this kind of problems is the {\em selective operator application} (SLA). In this case, each island deals with the whole chromosome, but only modifies a fragment of it in the crossover and mutation phase depending on the number of islands, using the whole chromosome for fitness calculation. This allows to deal with un-decomposable problems.
In our previous work \citep{Garcia16hpmoon} we used this method in a preliminary experimental setup, although it is worth mentioning we are establishing its name and acronym in this paper for the first time.  We proved
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. We discovered on that paper that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and synchronous island model. This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically adapts the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions. 
% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

The hypothesis of this paper is that the use of automatically calculated overlapped sections of the
chromosome for the SLA, can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. 
These overlapped sections are calculated from the number of available islands, in a
multi-objective island-based algorithm, and tested in a real computational environment.

%That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ FERGU: no, this is the hypothesis of this paper (comparing adaptive with partly overlapped), but clarifying it
In order to prove this claim, our new adaptive method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ FERGU: explained before that we are showing new results.
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ FERGU: Already writed in previous paragraph

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\journal{Applied Soft Computing}
\title{Distributed multi-objective evolutionary optimization using island-based selective operator application}

@@

%Introduction should start in the same way as the abstract. And
%starting with something that is not really applicable here, talking
%about general parallel evolutionary algorithms, is even worse.
% Right way (IMNSHO)
% 1. Multiobjective problems are
% 2. They usually require a good amount of computation → parallel
% distributed
% 3. One way to solve them is to use MOEAs _because_ they can be
% easily parallelized
% 4. But the easy paralellization is not necessarily efficient.
% 5. We have proposed the HPMOON (or whatever) technique to improve
% it. In paper (whatever) we proved that (whatever)
% 6. In this paper we will be proving whatever - JJ FERGU: Addressing JJ order
% OK - JJ

%1. MOP ARE
Multi-objective optimization problems (MOP) are those where several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest, considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives; the objective of MOP algorithms is to find as many
solutions in the Pareto Front as possible, and in order to do that the
exploration of the search space must be very efficient, trying to
cover many possible exploration paths at the same time.

%2 COMPUTATION -> PARALLEL
This aspect of searching for non-dominated solutions, as well as, in
many cases, the size of the search space itself, implies a high demand
of computational time, leading to the proposal of parallel and distributed methods to solve them
\citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268},
%3 ONE WAY IS TO USE MOEAS BECAUSE
one of which are evolutionary algorithms (EAs)
\citep{DBLP:series/ncs/EibenS15}. EAs are bio-inspired meta-heuristics
that can be effectively used to find nearly optimal solutions for
optimization problems and are easily paralellizable, since they are
population-based and creating a functionally equivalent parallel
version is as easy as distributing the population in communicating
computing nodes, which are usually called islands. The fact that
they are population-based helps find many solutions that compose the
Pareto fron simultaneously.

Usually an EA starts by generating a set of random solutions, called
\emph{population}, following a user-defined description. Then, it
evaluates each candidate solution, called \emph{individual}, assigning
it a \emph{fitness} value, that describes how good the individual is
with regard to the target problem. New solutions are then generated by
the application of \emph{operators} that either mutate a single
solution or recombine different existing solutions. After each
iteration, called \emph{generation}, the least fit individuals are
removed, and the process continues until a user-defined stop condition
is met.

One of the best known and referenced MOEAs
\citep{Dorronsoro13superlinear} is the {\em 
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, which are divided into different ranks, based on the
dominance, to be selected for the next generations. 
% This paragraph has nothing to do with anything. Is it useful for
% what you are doing in the paper? Is there any challenge that you
% will be solving here? For instance, you could talk about the problem
% it's got for parallelization and how it deals with them - JJ

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies to use different distribution
and sharing mechanisms, as there exist a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}. % Is that the case with NSGA? - JJ

MOPs usually require a high number of variables, % No they don't, and
                                % if they do, add a reference - JJ
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration. %depending on the
                                %language, this time might not be
                                %significant at all, and will depend
                                %on the length of the chromosome, not
                                %the part of the chromosome you will
                                %be applying it to. Besides, this is
                                %not the critical part of the
                                %algorithm, the critical part is
                                %searching the solution space. 
Different authors have % this is only one possibility. You should talk
                       % about simple island models, too - JJ
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a % first
                                % reference to co-evolution in this
                                % paper. Is it still a paper about
                                % co-evolution? - JJ
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

A similar approach used to solve this kind of problems is the {\em
  selective operator application} (SLA). In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. On that paper we proved that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically adapts the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions. 
% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

The hypothesis of this paper is that the use of automatically calculated overlapped sections of the
chromosome for the SLA can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. 
These overlapped sections are calculated from the number of available islands, in a
multi-objective island-based algorithm, and tested in a real computational environment.

In order to prove this claim, our new adaptive method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. 
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.

@@

\section*{References}
